Solving ...

### 2023-01-29 2023/PAVT_Ufa/PAVT2023.tex

The main part of the paper is organized as follows. Section 2 describes the mathematical model of the chemical reaction under study. The formal statement of the Lipschitz global optimization problem and the general scheme of search algorithms are given in Section 3. In Section 4, a scheme of the proposed asynchronous parallel algorithm for solving multiextremal problems is presented. The results of numerical solution of the inverse problem of chemical kinetics are discussed in Section 5.

@@

\subsection{Global Optimization Problem}

As it was mentioned above, the problem of identifying mathematical model parameter values can be considered a Lipschitz global optimization problem. From the formal point of view this problem is a mathematical programming problem of the form
\begin{gather}
 \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
\end{gather}
where $a,\; b$ are the given vectors, and $a,b\in R^N$, $\varphi(y)$ is the objective function which satisfies the Lipschitz condition
\begin{equation}\label{Lip}
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
\end{equation}

@@

There are several algorithms that can be applied to solve Lipschitz optimization problems. These include, among others, the non-uniform space covering method \cite{Evtushenko2013, Evtushenko2009}, diagonal and simplicial partitions methods \cite{Zilinskas2010, Paulavicius2011}. In this paper, we used the global search algorithm proposed by Strongin{Strongin2000}. Under this approach, the original multidimensional problem (\ref{problemN}) is reduced to a one-dimensional optimization problem using Peano-Hilbert curves.

@@

The Strongin algorithm has been constructed in the framework of the information approach to global optimization (see \cite{Strongin1978})). This method has an adaptive scheme to evaluate a numerical estimate of the unknown Lipschitz constant.

\subsection{Parallel Asynchronous Global Search Algorithm}

As the Global Search Algorithm (GSA) belongs to the class of characteristic algorithms, this suggests a possible way to parallelize it. 
As previously mentioned, the characteristic of the interval $R(i)$ can be regarded as some measure of finding the global minimum point in this interval. Then, instead of a single best interval, several intervals with the highest characteristics can be chosen at once and successive trials can be carried out in these intervals in parallel. Moreover, the scheme of characteristic algorithms also allows asynchronous parallelization, which minimizes downtime of processors when trial complexity depends on a particular point in the search domain.

@@

	\item
Assign to each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, the number $R(i)$, which is called the interval characteristic and is calculated using the formula
\[
R(i)=rM\Delta_i+\frac{(z_i-z_{i-1})^2}{rM\Delta_i}-2(z_i+z_{i-1}),
\]
where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.

@@

Numerical experiments were performed using the parallel asynchronous algorithm outlined in Section 4, running $160$ processes: $159$ processes calculated the values of the objective function (worker processes) while $1$ master process controlled the algorithm. The Lobachevsky  supercomputer (University of Nizhni Novgorod)  (CentOS 7.2, SLURM, two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM on the node) was used for the experiments. The asynchronous global optimization algorithm was implemented in C++ (GCC 9.5.0 and Intel MPI were used); the computation of objective function values was implemented in Python 3.9.

%Ускорение составило примерно $230$ раз.
The short time for solving the problem when using the parallel algorithm (approximately 14 minutes for 1 problem) allowed a thorough investigation of the dependence of the solution on the regularisation parameter $\alpha$. The table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter, the time for solving the problem (in seconds) and the time speedup in comparison to the estimated problem solution time when using the sequential method. The method parameter $r=4.0$ из (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$ were used in the runs. After a given number of iterations or after reaching the specified accuracy with the global search method, the solution was refined by the Hooke-Jeeves local method \cite{HookJeeves} with an accuracy of $\epsilon = 10^{-4}\left\|b-a\right\|$.

@@

\begin{table}
\caption{Исследование влияния параметра регуляризации alpha на минимальное значение функции}
\label{table_1}
\begin{center}
\begin{tabular}{cccc}
\hline\noalign{\smallskip}
 Alpha      & Minimum  & Time (sec.) & Speedup \\
\hline\noalign{\smallskip}
0.1		&	0.810454	&	806.192	&	241       \\
0.09	&	0.837868	&	828.649	&	234       \\
0.08	&	0.808238	&	811.852	&	239       \\
0.07	&	0.759134	&	803.626	&	242       \\
0.06	&	0.723404	&	810.9	&	240       \\
0.05	&	0.729415	&	804.674	&	242       \\
0.04	&	0.736035	&	816.387	&	238       \\
0.03	&	0.68152		&	809.185	&	240       \\
0.02	&	0.669092	&	155.837	&	3         \\
0.01	&	0.632944	&	825.238	&	235       \\
0.009	&	0.63001		&	814.48	&	239       \\
0.008	&	0.604817	&	830.448	&	234       \\
0.007	&	0.619546	&	807.335	&	241       \\
0.006	&	0.612614	&	834.703	&	233       \\
0.005	&	0.615638	&	815.45	&	238       \\
0.004	&	0.579725	&	821.421	&	237       \\
0.003	&	0.583529	&	831.454	&	234       \\
0.002	&	0.547907	&	850.017	&	229       \\
0.001	&	0.54075		&	853.898	&	227       \\
\noalign{\smallskip}\hline
\end{tabular}\end{center}\end{table}

>>>>

The main part of the paper is organized as follows. Section \ref{Sec_math_mod} describes the mathematical model of the chemical reaction under study. The formal statement of the Lipschitz global optimization problem and the general scheme of search algorithms are given in Section \ref{Sec_GO_Problem}. In Section \ref{Sec_PGSA}, a scheme of the proposed asynchronous parallel algorithm for solving multiextremal problems is presented. The results of numerical solution of the inverse problem of chemical kinetics are discussed in Section \ref{Sec_Exp}.

@@

\subsection{Global Optimization Problem}\label{Sec_GO_Problem}

As it was mentioned above, the problem of identifying mathematical model parameter values can be considered a Lipschitz global optimization problem. From the formal point of view this problem is a mathematical programming problem of the form
\begin{gather}
 \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
\end{gather}
where $a,b\in R^N$ are the given vectors and $\varphi(y)$ is the objective function which satisfies the Lipschitz condition
\begin{equation}\label{Lip}
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
\end{equation}

@@

There are several algorithms that can be applied to solve Lipschitz optimization problems. These include, among others, the non-uniform space covering method \cite{Evtushenko2013, Evtushenko2009}, diagonal and simplicial partitions methods \cite{Zilinskas2010, Paulavicius2011}. In this paper, we used the global search algorithm proposed by Strongin\cite{Strongin2000}. Under this approach, the original multidimensional problem (\ref{problemN}) is reduced to a one-dimensional optimization problem using Peano-Hilbert curves.

@@

The Strongin global search algorithm has been constructed in the framework of the information approach to global optimization (see \cite{Strongin2000})). This method has an adaptive scheme to evaluate a numerical estimate of the unknown Lipschitz constant.

\subsection{Parallel Asynchronous Global Search Algorithm}\label{Sec_PGSA}

As the global search algorithm (GSA) belongs to the class of characteristic algorithms, this suggests a possible way to parallelize it. 
As previously mentioned, the characteristic of the interval $R(i)$ can be regarded as some measure of finding the global minimum point in this interval. Then, instead of a single best interval, several intervals with the highest characteristics can be chosen at once and successive trials can be carried out in these intervals in parallel. Moreover, the scheme of characteristic algorithms also allows asynchronous parallelization, which minimizes downtime of processors when trial complexity depends on a particular point in the search domain.

@@

	\item
Assign to each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, the number $R(i)$, which is called the interval characteristic and is calculated using the formula
\begin{equation} \label{R}
R(i)=rM\Delta_i+\frac{(z_i-z_{i-1})^2}{rM\Delta_i}-2(z_i+z_{i-1}),
\end{equation}
where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.

@@

Numerical experiments were performed using the parallel asynchronous algorithm outlined in Section \ref{Sec_PGSA}, running $160$ processes: $159$ processes calculated the values of the objective function (worker processes) while $1$ master process controlled the algorithm. The Lobachevsky  supercomputer (University of Nizhni Novgorod)  (CentOS 7.2, SLURM, two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM on the node) was used for the experiments. The asynchronous global optimization algorithm was implemented in C++ (GCC 9.5.0 and Intel MPI were used); the computation of objective function values was implemented in Python 3.9.

%Ускорение составило примерно $230$ раз.
The short time for solving the problem when using the parallel algorithm (approximately 14 minutes for 1 problem) allowed a thorough investigation of the dependence of the solution on the regularisation parameter $\alpha$. The table \ref{table_1} shows the minimum values of the objective function found for the corresponding value of the regularization parameter, the time for solving the problem (in seconds) and the time speedup in comparison to the estimated problem solution time when using the sequential method. The method parameter $r=4.0$ from (\ref{R}) and the accuracy $\epsilon = 10^{-4}\left\|b-a\right\|$ were used in the runs. After a given number of iterations or after reaching the specified accuracy with the global search method, the solution was refined by the Hooke-Jeeves local method \cite{HookJeeves} with an accuracy of $\epsilon = 10^{-4}\left\|b-a\right\|$.

@@

\begin{table}
\caption{Dependence of the function minimum value $\varphi^*$ on the regularization parameter $\alpha$}
\label{table_1}
\begin{center}
\begin{tabular}{ccc}
\hline\noalign{\smallskip}
 $\alpha$      & $\varphi^*$  & Time (sec.) \\
\hline\noalign{\smallskip}
0.1		&	0.810454	&	806.192	\\
0.09	&	0.837868	&	828.649 \\
0.08	&	0.808238	&	811.852	\\
0.07	&	0.759134	&	803.626 \\
0.06	&	0.723404	&	810.9		\\
0.05	&	0.729415	&	804.674	\\
0.04	&	0.736035	&	816.387	\\
0.03	&	0.68152		&	809.185	\\
0.02	&	0.669092	&	155.837	\\
0.01	&	0.632944	&	825.238	\\
0.009	&	0.63001		&	814.48	\\
0.008	&	0.604817	&	830.448	\\
0.007	&	0.619546	&	807.335	\\
0.006	&	0.612614	&	834.703	\\
0.005	&	0.615638	&	815.45	\\
0.004	&	0.579725	&	821.421	\\
0.003	&	0.583529	&	831.454	\\
0.002	&	0.547907	&	850.017	\\
0.001	&	0.54075		&	853.898	\\
\noalign{\smallskip}\hline
\end{tabular}\end{center}\end{table}
